\replaced{Towards an o}{O}pen-source landscape for 3D CSEM modelling

### 2021-01-28 manuscript/Manuscript.tex

\let\tablenum\relax
\usepackage{siunitx}
\usepackage{tabularx}

@@

% Own commands
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\emg}[2]{\texttt{emg#1#2}\xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\simpeg}{\texttt{SimPEG}\xspace}
\newcommand{\discretize}{\texttt{discretize}\xspace}
\newcommand{\custem}{\texttt{custEM}\xspace}
\newcommand{\petgem}{\texttt{PETGEM}\xspace}

% Remove once done
% \newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
% \newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
% \newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
\newcommand{\dul}[1]{\underline{\underline{#1}}}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\rmk}[1]{{\color{red}{#1}}}
\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}

\title[3D CSEM Open-Source Landscape]{\replaced{Towards an open-source landscape for 3D CSEM modelling}{Open-source landscape for 3D CSEM modelling}}

@@

\date{Received 202Y MMMM DD; in original form 2020 October 23}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{VVV}
\pubyear{202Y}

@@

@@

% TODO delete before re-submission
\begin{itemize}
  \item \deleted{old}: \texttt{\textbackslash deleted\{old\}}
  \item \added{new}: \texttt{\textbackslash added\{new\}}
  \item \replaced{new}{old}: \texttt{\textbackslash replaced\{new\}\{old\}}
\end{itemize}

@@

Many of the advancements made in the EM modelling community over the past decades have required that authors develop new implementations from scratch. These codes often provided the research group or company with a competitive advantage for a time, and thus the source codes were often kept internal. In some cases, executables have been made available for academic purposes upon request or to sponsoring consortium members. As the field continues to mature, advancements become more incremental, and particularly in an applied field, many advancements are driven by new use-cases and applications that were not considered by the original authors. In the aforementioned review on EM modelling and inversion Dmitry Avdeev concludes with the following statement: «\emph{The most important challenge that faces the EM community today is to convince software developers to put their 3-D EM forward and inverse solutions into the public domain, at least after some time. This would have a strong impact on the whole subject and the developers would benefit from feedback regarding the real needs of the end-users.}» Similarly, \cite{EXG.19.Oldenburg} argue that an open-source paradigm has the potential to accelerate multidisciplinary research by encouraging the development of modular, interoperable tools that are supported by a community of researchers.

@@

The open-source landscape in the field of 3D CSEM modelling changed in the last five years quite dramatically, and here we introduce and compare four recent open-source projects. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes. Having different codes which use different methods and different meshes is ideal to address the topic of validation. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. The only objective possibility to ensure the accuracy of solutions beyond these simple models is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of MT forward and inversion codes, but we are not aware of any comparable comparison study or benchmark suite for CSEM data.

We simulate EM fields for a layered background model with vertical transverse isotropy, with and without three resistive blocks, as well as for the complex marine, open-source MR3D model. These three models as well as the corresponding results from four different codes provide a benchmark for new (and existing) codes to be compared to and validated with. First we introduce the codes under consideration, and present afterwards the considered benchmark cases in detail together with the modelling results of our four codes in terms of accuracy and computational performance. Beyond that, we extensively discuss important points that control the performance and suitability of our FV and FE codes, including considerations about the mesh design and the choice of solvers. We conclude with a discussion and conclusions, and a motivation for the EM community at large to not only continue to extend the landscape of open-source codes but to also create a landscape of open-source benchmark models.

@@

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing routines written in Python, and all of them make heavy use of \texttt{NumPy} \citep{NAT.20.Harris} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four of them follow the open model of development, meaning that they come with both\deleted{,} an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests). All developments comprise an extensive online documentation with many examples and have continuous integration to some degree. Newer package-management systems such as \texttt{conda}, \texttt{docker}, or \texttt{pip} simplify installation for all of these codes on any major operating system

We briefly present our codes in this section. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the codes is given in Table~\ref{tbl:codecomparison}. All four have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range usually applied in CSEM. \added{For the shown numerical examples all codes use the total-field formulation for the electric field.} The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator. A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations \emph{PEC} and \emph{PMC}, which stand for perfect electric conductor
% We could add maths here; but we would need to introduce all the symbols. For now, we leave it out.
% ($\vec{n}\times\vec{E} = 0; \vec{n}\cdot\vec{H} = 0$)
and perfect magnetic conductor%
% ($\vec{n}\cdot\vec{E} = 0; \vec{n}\times\vec{H} = 0$)
, respectively. Another abbreviation used in the tables is \emph{dof} for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real-world time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation time or CPU time is therefore much higher than the reported runtime for the codes that run in parallel, with an upper limit of \#Procs$\times$Runtime. Memory refers to the maximum memory increase at any point of solving the SLE.

%
\begin{table*}
\begin{minipage}{.9\linewidth}
  \centering
  \caption{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}
\label{tbl:codecomparison}
  \begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Website & \href{https://custem.rtfd.io}{custem.rtfd.io}
          & \href{https://empymod.github.io}{empymod.github.io}
          & \href{http://petgem.bsc.es}{petgem.bsc.es}
          & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & BSD-3-Clause & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti-/curvilinear, octree \\
  BC & PEC; PMC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}
\end{minipage}
\end{table*}
%

%
\begin{table*}
\begin{minipage}{14cm}
  \centering
  \caption{List of hardware, software, and operator with which the different codes were run.}
\label{tbl:machines}
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 (server)}                  & Raphael Rochlitz \\
          & - 144 Xeon Gold 6154 CPU @2.666 GHz               & \\
          & - $\approx3$ TB DDR4 RAM                          & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \emg3d  & \textbf{Dell Latitude (laptop)}                   & Dieter Werthmüller \\
          & - i7-6600U CPU@2.6 GHz x4                         & \\
          & - 15.5 GiB RAM                                    & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \petgem & \textbf{Marenostrum4 (supercomputer)}             & Octavio Castillo-Reyes \\
          & - 2 sockets Intel Xeon Platinum (Skylake generation) 8160 CPU & \\
          & \phantom{-} with 24 cores each @2.10GHz for a total of 48 cores per node & \\
          & - 386 GB DDR4 RAM per node                        & \\
          & - SuSE Linux Enterprise                           & \\[.5em]
  \simpeg & \textbf{GKE n2-custom (Google cloud)}             & Lindsey Heagy \\
          & - Intel Cascade Lake, 8 vCPUs                     & \\
          & - 355 GB RAM                                      & \\
          & - Ubuntu 16.04                                    & \\
%
  \bottomrule
\end{tabularx}
\end{minipage}
\end{table*}
%

@@

The toolbox considers generally anisotropic petrophysical properties. Even though changes of the conductivity are mainly of interest for CSEM modelling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements. Recently, induced polarisation parameters in frequency-domain and three methods for simulating time-domain responses were added to \custem. The provided meshing tools are based on \texttt{TetGen} \citep{TOM.15.Si} and functionalities of \texttt{pyGIMLi} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

@@

As such the code is primarily a solver, which solves Maxwell's equations for a single frequency and a single, electric source. Recent developments added functionalities such that \emg3d can be used directly as a more general EM modeller too. It features now routines to obtain the magnetic field due to an electric source, to obtain the electric and magnetic fields due to a magnetic source, and to obtain time-domain responses. For the underlying discretization the Python package \discretize is used, which is part of the larger \simpeg ecosystem.

@@

For the parallel forward modelling computations, a highly scalable MPI (message passing interface) domain decomposition allows reducing runtimes and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters (e.g., High-Performance Computing [HPC] architectures).

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in geophysics. Current functionality includes modelling and inversion capabilities for gravity, magnetics, direct current resistivity, induced polarisation, EM (time and frequency, as well as natural sources), and fluid flow. It is a community driven project that aims to support researchers and practitioners by providing a flexible, extensible framework, and common interface to a variety of geophysical methods.

@@

We computed the responses for three different models to validate that the four 3D CSEM codes yield the same electromagnetic responses and compare different solver types (FE, FV) and different mesh types (unstructured tetrahedra, rectilinear, octree) in different scenarios. The first model is an anisotropic layered model, which can be compared to semi-analytical solutions. The layered model also serves as background model for the second validation, where we add three resistive blocks into the subsurface. The final comparison is based on the realistic, anisotropic marine model MR3D. It is an open resistivity model which comes with computed CSEM responses.

@@

A layered model fails to show the strength of 3D modellers, and in reality one would not choose a 3D code to compute responses for a layered model. However, it is one of the few examples that can be verified with semi-analytical results, and is therefore still a valuable test. The results are shown in Figure~\ref{fig:results-layered}, where the left, middle, and right columns are for the real part, imaginary part, and absolute value, respectively (dashed lines indicate negative values). The top row shows the actual, semi-analytically computed responses, for which we used \empymod\citep{GEO.17.Werthmuller}. The second and the third row show the relative percentage error of the inline and the broadside receivers, respectively; note that for the 1D case the two broadside lines for $y=\pm3\,$km are identical.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-layered}
  \caption{Results of the layered model: The semi-analytical inline and broadside responses are shown in the top row. The relative percentage error of the four 3D modellers are shown in the second and the third row for the inline and broadside responses, respectively.}
  \label{fig:results-layered}
\end{figure*}
%

There are a few noteworthy points:
\begin{enumerate}
  \item The error of the real and imaginary responses are dominated by their zero-crossings, which blows up the error as is the nature of the relative error formula, it has no real-world implications.
  \item It can be seen from the absolute-value errors that the misfit is generally in the order of 1\,\% or less.
  \item The errors become substantial towards zero offset, particularly for the results of the FV codes. These comparably significant errors could be significantly reduced by choosing a finer discretization around the source, at the cost of increased computation time.
  \item The error of the responses from the FE codes, on the hand, are relatively big at large offsets, particularly for the inline real responses. This has, again, its origin in the chosen discretization. The thin isotropic layer beneath the water requires many tetrahedra for a good-quality discretization. Therefore, the FE codes only extend this thin layer up to 30\,km from the origin, surrounded by a ten times larger halfspace-like boundary mesh with water conductivities assigned to the subsurface. This approximation significantly reduces the problem size compared to considering the layered-earth geometry for the whole computational domain.
\end{enumerate}

@@

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}
\label{tbl:comp-layered}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 24 &   118 &  97.8 & 1531824 \\ % dof: 3063648 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   116 &   0.4 & 4813216 \\
  \petgem & 24 &   252 & 151.7 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

The block model is a derivation of the \emph{Dublin Test Model 1} from the first EM modelling workshop described by \cite{GJI.13.Miensopust}. We use the same layout of the blocks but adjust the dimensions and resistivities to a typical marine CSEM problem, as shown in Figure~\ref{fig:model-block}. Additionally, we add our \emph{Layered Model} as a VTI background. The three resistive blocks have resistivities of $\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines), $\rho=100\,\ohmm$ (thin plate, South-East), and $\rho=500\,\ohmm$ (cube, North-West).
%
\begin{figure*}
  \centering
  \includegraphics[width=.6\linewidth]{figures/model-block}
  \caption{Sketch of the block model, consisting of the layered background model with three resistive blocks, embedded in the thick background layer which has VTI with $\lambda=\sqrt{\rho_\textrm{v}/\rho_\textrm{h}}=\sqrt{2}$.}
  \label{fig:model-block}
\end{figure*}
%

@@

The results for the three receiver lines $y=-3;0;+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the normalised differences between the absolute responses of the codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-block}
  \caption{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the normalised differences (\%) between the absolute responses $|E_x|$ of the different codes are shown in the bottom row.}
  \label{fig:results-block}
\end{figure*}
%

@@

@@

The shown examples consider the geophysical problem of marine CSEM with resistive bodies in the frequency domain. We hope to see in the future similar comparisons for other cases, such as land CSEM with strong topography; airborne and mixed air-ground surveys; looking for conductive bodies; time domain; and models based on horizons instead of blocks which pose a challenge for structured grids and favour unstructured meshes.

@@

\section{Data}

>>>>

\let\tablenum\relax
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage{tcolorbox}

@@

% Own commands
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\emg}[2]{\texttt{emg#1#2}\xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\simpeg}{\texttt{SimPEG}\xspace}
\newcommand{\discretize}{\texttt{discretize}\xspace}
\newcommand{\custem}{\texttt{custEM}\xspace}
\newcommand{\petgem}{\texttt{PETGEM}\xspace}
\newcommand{\ymark}[1]{%
  \begin{tcolorbox}[colback=yellow!40!white]#1\end{tcolorbox}}

% Remove once done
\newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
% \newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
\newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
\newcommand{\dul}[1]{\underline{\underline{#1}}}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\rmk}[1]{{\color{red}{#1}}}
\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}

\title[3D CSEM Open-Source Landscape]{\replaced{Towards an o}{O}pen-source landscape for 3D CSEM modelling}

@@

\date{Received 2021 MMMM DD; in original form 2020 October 23}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{VVV}
\pubyear{202Y}

@@

\clearpage

\subsection*{Correction commands}
\begin{itemize}
  \item \deleted{old}: \texttt{\textbackslash deleted\{old\}}
  \item \added{new}: \texttt{\textbackslash added\{new\}}
  \item \replaced{new}{old}: \texttt{\textbackslash replaced\{new\}\{old\}}
\end{itemize}

\subsection*{Revise entire manuscript re:}
\ymark{Intro/Discussion: Take into account «To make this manuscript publishable within GJI, it requires some serious effort to focus more on the bigger picture and make this story more pronounced within your manuscript.»}

\listoftodos

\clearpage

@@

@@

Many of the advancements made in the EM modelling community over the past decades have required that authors develop new implementations from scratch. These codes often provided the research group or company with a competitive advantage for a time, and thus the source codes were \replaced{normally}{often} kept internal. In some cases, executables have been made available for academic purposes upon request or to sponsoring consortium members. As the field continues to mature, advancements become more incremental, and particularly in an applied field, many advancements are driven by new use-cases and applications that were not considered by the original authors. In the aforementioned review on EM modelling and inversion Dmitry Avdeev concludes with the following statement: «\emph{The most important challenge that faces the EM community today is to convince software developers to put their 3-D EM forward and inverse solutions into the public domain, at least after some time. This would have a strong impact on the whole subject and the developers would benefit from feedback regarding the real needs of the end-users.}» Similarly, \cite{EXG.19.Oldenburg} argue that an open-source paradigm has the potential to accelerate multidisciplinary research by encouraging the development of modular, interoperable tools that are supported by a community of researchers.

@@

The landscape in the field of 3D CSEM modelling changed in the last five years quite dramatically. This paper introduces and compares four projects which make all source code openly available for use and adaptation. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes. Having different codes which use different methods and different meshes is ideal to address the topic of validation, and having all openly available facilitates reproducibility of our results. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. The only objective possibility to ensure the accuracy of solutions beyond these simple models is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of MT forward and inversion codes, but we are not aware of any comparable comparison study or benchmark suite for CSEM data.

We simulate EM fields for a layered background model with vertical transverse isotropy, with and without three resistive blocks, as well as for the complex marine, open-source MR3D model. These three models as well as the corresponding results from four different codes provide a benchmark for new (and existing) codes to be compared to and validated with. First we introduce the codes under consideration, and present afterwards the considered benchmark cases in detail together with the modelling results of our four codes in terms of accuracy and computational performance. Beyond that, we extensively discuss important points that control the performance and suitability of our FV and FE codes, including considerations about the mesh design and the choice of solvers. Such a comparison reveals avenues for further development for each of the codes. We conclude with a discussion and conclusions, and a motivation for the EM community at large to not only continue to extend the landscape of open-source codes but to also create a landscape of open-source benchmark models.

@@

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing routines written in Python, and all of them make heavy use of \texttt{NumPy} \citep{NAT.20.Harris} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four of them follow the open model of development, meaning that they come with both\deleted{,} an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests). All developments comprise an extensive online documentation with many examples and have continuous integration to some degree. Newer package-management systems such as \texttt{conda}, \texttt{docker}, or \texttt{pip} simplify installation for all of these codes on any major operating system.

We briefly present our codes in this section. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the codes is given in Table~\ref{tbl:codecomparison}. All four have in common that they solve Maxwell's equation in its differential form \deleted{(DE)} under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range usually applied in CSEM. \added{For the numerical examples we show, all codes use the total-field formulation for the electric field.} The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator. A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations \emph{PEC} and \emph{PMC}, which stand for perfect electric conductor
% We could add maths here; but we would need to introduce all the symbols. For now, we leave it out.
% ($\vec{n}\times\vec{E} = 0; \vec{n}\cdot\vec{H} = 0$)
and perfect magnetic conductor%
% ($\vec{n}\cdot\vec{E} = 0; \vec{n}\times\vec{H} = 0$)
, respectively. \added{PEC (PMC) implies that the tangential electric (magnetic) field vector components are zero at the boundary.} Another abbreviation used in the tables is \emph{dof} for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real-world time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation time or CPU time is therefore much higher than the reported runtime for the codes that run in parallel, with an upper limit of \#Procs$\times$Runtime. Memory refers to the maximum memory increase at any point of solving the SLE.

%
\begin{table*}
\begin{minipage}{.9\linewidth}
  \centering
  \caption{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}
\label{tbl:codecomparison}
  \begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Website & \href{https://custem.rtfd.io}{custem.rtfd.io}
          & \href{https://emsig.github.io}{emsig.github.io}
          & \href{http://petgem.bsc.es}{petgem.bsc.es}
          & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & BSD-3-Clause & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti-/curvilinear, octree \\
  BC & PEC; PMC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}
\end{minipage}
\end{table*}
%

%
\begin{table*}
\begin{minipage}{14cm}
  \centering
  \caption{List of hardware, software, and operator with which the different codes were run.}
\label{tbl:machines}
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 (server)}                  & Raphael Rochlitz \\
          & - 144 Xeon Gold 6154 CPU @2.666 GHz               & \\
          & - $\approx3$ TB DDR4 RAM                          & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \emg3d  & \textbf{Dell Latitude (laptop)}                   & Dieter Werthmüller \\
          & - i7-6600U CPU@2.6 GHz x4                         & \\
          & - 15.5 GiB RAM                                    & \\
          & - Ubuntu 20.04                                    & \\[.5em]
  \petgem & \textbf{Marenostrum4 (supercomputer)}             & Octavio Castillo-Reyes \\
          & - 2 sockets Intel Xeon Platinum (Skylake generation) 8160 CPU & \\
          & \phantom{-} with 24 cores each @2.10GHz for a total of 48 cores per node & \\
          & - 386 GB DDR4 RAM per node                        & \\
          & - SuSE Linux Enterprise                           & \\[.5em]
  \simpeg & \textbf{GKE n2-custom (Google cloud)}             & Lindsey Heagy \\
          & - Intel Cascade Lake, 8 vCPUs                     & \\
          & - 355 GB RAM                                      & \\
          & - Ubuntu 16.04                                    & \\
%
  \bottomrule
\end{tabularx}
\end{minipage}
\end{table*}
%

@@

The toolbox considers generally anisotropic petrophysical properties. Even though changes of the conductivity are mainly of interest for CSEM modelling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements. Recently, induced polarisation parameters in frequency-domain \added{computations} and three methods for simulating time-domain responses were added to \custem. The provided meshing tools are based on \texttt{TetGen} \citep{TOM.15.Si} and functionalities of \texttt{pyGIMLi} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

@@

As such the code is primarily a solver, which solves Maxwell's equations for a single frequency and a single, electric source. Recent developments added functionalities such that \emg3d can be used directly as a more general EM modeller too. It features now \replaced{besides arbitrarily rotated electric sources and receivers also routines for magnetic sources and receivers, to model time-domain responses, and to obtain the gradient of the misfit function which can be used by inversion algorithms}{routines to obtain the magnetic field due to an electric source, to obtain the electric and magnetic fields due to a magnetic source, and to obtain time-domain responses}. For the underlying discretization the Python package \discretize is used, which is part of the larger \simpeg ecosystem.

@@

For the parallel forward modelling computations, a highly scalable MPI (message passing interface) domain decomposition allows reducing runtimes and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters \replaced{such as}{(e.g.,} High-Performance Computing (HPC) architectures\deleted{)}.

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in geophysics. Current functionality includes modelling and inversion capabilities for gravity, magnetics, direct current resistivity, induced polarisation, EM (time and frequency \added{domain}, \replaced{controlled and}{as well as} natural sources), and fluid flow. It is a community driven project that aims to support researchers and practitioners by providing a flexible, extensible framework, and common interface to a variety of geophysical methods.

@@

We computed the responses for three different models to validate that the four 3D CSEM codes yield the same electromagnetic responses and compare different solver types (FE, FV) and different mesh types (unstructured tetrahedra, rectilinear, octree) in different scenarios. The first model is an anisotropic\added{,} layered model, which can be compared to semi-analytical solutions. The layered model also serves as background model for the second validation, where we add three resistive blocks into the subsurface. The final comparison is based on the realistic, anisotropic marine model MR3D. It is a\deleted{n open} resistivity model \replaced{that is openly released, together}{which comes} with computed CSEM responses.

@@

A layered model fails to show the strength of 3D modellers, and in reality one would not choose a 3D code to compute responses for a layered model. However, it is one of the few examples that can be verified with semi-analytical results, and is therefore still a valuable test. \itodo{Adjust everything here in section 3.1 for new figures}The results are shown in Figure~\ref{fig:results-layered}, where the left, middle, and right columns are for the real part, imaginary part, and absolute value, respectively (dashed lines indicate negative values). The top row shows the actual, semi-analytically computed responses, for which we used \empymod\citep{GEO.17.Werthmuller}. The second and the third row show the relative percentage error of the inline and the broadside receivers, respectively; note that for the 1D case the two broadside lines for $y=\pm3\,$km are identical.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-layered}
  \caption{\itodo{Adjust Figures 1 (Amplitude/Phase) and corresponding text in manuscript.} Results of the layered model: The semi-analytical inline and broadside responses are shown in the top row. The relative percentage error of the four 3D modellers are shown in the second and the third row for the inline and broadside responses, respectively.}
  \label{fig:results-layered}
\end{figure*}
%

There are a few noteworthy points:
\begin{enumerate}
  \item \itodo{adjust}The error of the real and imaginary responses are dominated by their zero-crossings, which blows up the error as is the nature of the relative error formula, it has no real-world implications.
  \item It can be seen from the absolute-value errors that the misfit is generally in the order of 1\,\% or less.
  \item The errors become substantial towards zero offset, particularly for the results of the FV codes. These comparably significant errors could be significantly reduced by choosing a finer discretization around the source, at the cost of increased computation time. \itodo{Add a sentence re close offsets FV codes, Rune's comment (or above in point ii).}
  \item~\\ \itodo{Adjust re changes of Raphael.} The error of the responses from the FE codes, on the hand, are relatively big at large offsets, particularly for the inline real responses. This has, again, its origin in the chosen discretization. The thin isotropic layer beneath the water requires many tetrahedra for a good-quality discretization. Therefore, the FE codes only extend this thin layer up to 30\,km from the origin, surrounded by a ten times larger halfspace-like boundary mesh with water conductivities assigned to the subsurface. This approximation significantly reduces the problem size compared to considering the layered-earth geometry for the whole computational domain.
\end{enumerate}

@@

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{\itodo{Check all}Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}
\label{tbl:comp-layered}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 24 &   118 &  97.8 & 1531824 \\ % dof: 3063648 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   116 &   0.4 & 4813216 \\
  \petgem & 24 &   252 & 151.7 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

\itodo{Explain/clarify re interpolation and re vertical fields in the manuscript.}

@@

The block model is a derivation of the \emph{Dublin Test Model~1} from the first EM modelling workshop described by \cite{GJI.13.Miensopust}. We use the same layout of the blocks but adjust the dimensions and resistivities to a typical marine CSEM problem, as shown in Figure~\ref{fig:model-block}. Additionally, we add our \emph{Layered Model} as a VTI background. The three resistive blocks have resistivities of $\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines), $\rho=100\,\ohmm$ (thin plate, South-East), and $\rho=500\,\ohmm$ (cube, North-West).
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/model-block}
  \caption{Sketch of the block model, consisting of the layered background model with three resistive blocks, embedded in the thick background layer which has VTI with $\lambda=\sqrt{\rho_\textrm{v}/\rho_\textrm{h}}=\sqrt{2}$.}
  \label{fig:model-block}
\end{figure}
%

@@

The results for the three receiver lines $y=-3;0;+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the normalised differences between the absolute responses of the codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-block}
  \caption{\itodo{Adjust Figures 3 (\emg3d-\simpeg; \custem-\petgem; FV-FE) and corresponding text in manuscript.}Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the normalised differences (\%) between the absolute responses $|E_x|$ of the different codes are shown in the bottom row.}
  \label{fig:results-block}
\end{figure*}
%

@@

\itodo{SOMEWHERE: Address NRMSD of MR3D results better; EMGS code elan uses bathymetry file; difference not error; very strong NRMSD; although visually identical.}

@@

The shown examples consider the geophysical problem of marine CSEM with resistive bodies in the frequency domain. \added{Marlim R3D is a rare example of an open-source resistivity model that also comes with CSEM data. The spectrum of geophysical EM methods is much wider than marine CSEM.} We hope to see in the future \added{more open-source models with accompanying data and} similar comparisons for other cases, such as land CSEM with strong topography; airborne and mixed air-ground surveys; looking for conductive bodies; time domain; and models based on horizons instead of blocks which pose a challenge for structured grids and favour unstructured meshes. \added{An ideal realistic model should be defined as a function resistivity(x, y, z) which returns the resistivity for any location in the domain of interest.}

@@

\section{Data}
\itodo{Zenodo DOI: Upload and link here.}
The resulting CSEM responses of the four codes for the three models, as well as all files to rerun the different models with the four codes and reproduce the shown results are available at Zenodo (DOI) \emph{with the final submission}. For the review process we refer to the github page \href{https://github.com/swung-research/3d-csem-open-source-landscape}{github.com/swung-research/3d-csem-open-source-landscape}.
